An Embedding Framework for Consistent Polyhedral Surrogates

### 2021-06-18 poly-embed-journal/poly-embed-journal.tex
sec 6 start; probably done with reorg up

\subsection{Structure of polyhedral Bayes risks}
\raf{Lemma X}

@@

\begin{proposition}\label{prop:SL-minimum}
  A minimizable loss $L: \R \to \reals^\Y_+$ has a polyhedral Bayes risk if and only if it has a finite representive set.
  % Moreover, for any minimum representative set $\Sc$ for $L$, there is a bijection between $\Sc$ and the facets of the epigraph of $-\risk{L}$.
  % \raft{Jessie: prove :-].  Note: min rep $\implies$ poly risk is immediate from Lemma~\ref{lem:loss-restrict}.  Move the construction into the proof. \jessie{Done; ready for a look}}
\end{proposition}
\begin{proof}
  A finite representative implies a polyhedral Bayes risk by Lemma~\ref{lem:loss-restrict}.

  For the converse, suppose $L$ has a polyhedral Bayes risk $\risk{L}$.
  Then we may write $\risk{L}(p) = \min_{(a,b)}$

  Let $\F$ be the set of facets of the epigraph of the negative Bayes risk

  implies it has a finite minimum representative set, we construct the set $\Sc$ as follows:

  since the epigraph of negative Bayes risk $E_L$ is polyhedral, it has finitely many facets, and we iterate over these.
  Each facet $F$ of $E_L$ must be full dimensional by definition of a facet, and therefore there must be some $p \in \inter{F}$.
  Observe that $p$ must by uniquely supported by some hyperplane $p \mapsto (p, L(u))$ for some $u \in \R$, since it is in the interior of the facet (and therefore $\risk{L}$ is differentiable at $p$).
  Add $u$ to $\Sc$, and repeat, iterating over each of $E_L$'s finitely many facets.
  We now show $\Sc$ is a minimum representative set.

  First, we have to show that $\Sc$ is a representative set.
  As facets are closed, for each facet $F$ and $p_F \in F$, there is a report $u_F \in \Sc$ by construction and closedness of facets so that the hyperplane $p_F \mapsto (p_F, \inprod{p_F}{L(u_F)})$ supports $E_L$ exactly on the facet. Therefore $u_F \in \Gamma(p) \cap \Sc$ for all $p$ such that the hyperplane supports $E_L$ at $p$.
  Representativeness follows as the facets of $E_L$ union to the simplex.

  If $\Sc$ was not a \emph{minimum} representative set \jessiet{minimal proof, not minimum}, then we could remove some $s$ from $\Sc$ such that $\Sc' := \Sc \setminus \{s\}$ is also a representative set for $L$.
  By the construction of $\Sc$ over facets of $E_L$, this report $s$ corresponds to a facet $F$ of $E_L$ since facets are full-dimensional and thus have well-defined nonempty interiors.
  There is then some distribution $p_F \in \inter F$ such that the hyperplane $p \mapsto (p, L(s))$ uniquely supports $E_L$ at $p_F$.
  We claim $\Gamma(p) \cap \Sc' = \emptyset$, since any other element $s'$ of $\Sc'$ being in $\Gamma(p)$ would imply $p \mapsto (p, L(s'))$ also supports $E_L$, but this hyperplane supporting $E_L$ must be unique, so these are equal.
  By construction of $\Sc$, we must have only added at most one of these to the set.
  Therefore, $\Sc$ is a minimum representative set.

  Moreover, we know there is a bijection between $\Sc$ and the facets of $E_L$ since $\card(\Sc) = |\Sc| = |\{F \mid F$ facet of $E_L\}| = \card($facets of $E_L)$ by construction of $\Sc$.
\end{proof}

@@

\begin{proposition}\label{prop:embed-bayes-risks}
  Let discrete loss $\ell$ and minimizable loss $L$ be given.
  Then $L$ embeds $\ell$ if and only if $\risk{L}=\risk{\ell}$.
  % Let discrete loss $\ell:\R\to\reals^Y$ be given.
  % Then $L:\reals^d\to\reals^\Y$ embeds $\ell$ if and only if $\risk{L}=\risk{\ell}$.
  % A loss $L$ embeds a discrete loss $\ell$ if and only if $\risk{L}=\risk{\ell}$.
\end{proposition}
\begin{proof}
  \raft{4/2/2021 first part of the proof should be done now.}
  Define $\Gamma = \prop{L}$ and $\gamma = \prop{\ell}$.
  Suppose $L$ embeds $\ell$, so we have some $\Sc\subseteq \R$ which is representative for $\ell$ and an embedding $\varphi:\Sc\to\reals^d$; take $\U := \varphi(\Sc)$.
  Since $\Sc$ is representative for $\ell$, by embedding condition (ii) we have $\{\gamma_s \mid s\in\Sc\} = \{\Gamma_u \mid u\in\U\}$, so $\U$ is representative for $L$.
  % for all $p\in\simplex$ we have some $r \in \Sc \cap \gamma(p)$ and thus we have $\varphi(r) \in \U \cap \Gamma(p)$.
  % In particular, $\U \cap \Gamma(p)\neq \emptyset$ for all $p\in\simplex$, 
  By Lemma~\ref{lem:loss-restrict}, we have $\risk{\ell} = \risk{\ell|_{\Sc}}$ and $\risk{L} = \risk{L|_{\U}}$.
  % \raft{Commented out: the symmetric argument that representation goes both ways under an embedding.  I'm thinking we should expand on this perspective in section 5, maybe introduce a notion of equivalence between any two losses (discrete or otherwise), and define trim in that context.}
  % Taking $\U := \varphi(\Sc)$, we have that $\ell$ if and only if $\U$ is representative of $L$, as $\gamma(p) \cap \Sc \neq \emptyset \iff \exists r \in \Sc$ such that $r \in \gamma(p) \iff \exists r$ such that $\varphi(r) \in \Gamma(p) \cap \U \iff \Gamma(p) \cap \U \neq \emptyset$.
%  To see that $\gamma'(p) \neq \emptyset$ for all $p\in\simplex$, note that by the definition of $\gamma$ as the property elicited by $\ell$ we have some $r \in \gamma(p)$, and by the embedding condition~\eqref{eq:embed-loss}, $\varphi(r) \in \Gamma(p)$.
  As $L(\varphi(\cdot)) = \ell(\cdot)$ by embedding condition (i), for all $p\in\simplex$ we have
  \begin{equation*}
    \risk{\ell}(p) = \risk{\ell|_\Sc}(p) = \min_{r \in \Sc}\inprod{p}{\ell(r)} = \min_{r \in \Sc}\inprod{p}{L(\varphi(r))} = \min_{u \in \U}\inprod{p}{L(u)} = \risk{L|_\U}(p) = \risk{L}(p)~.
  \end{equation*}

    %\btw{RF: This is on the informal side, mostly because of the ``working in the affine hull'' bit.  It can all be formalized though, probably should be for the journal version: just take an appropriate linear projection of $\reals^n$ to $\reals^{n-1}$, and work with the projected versions of everything, and then project back.}

% commented out 10 May 21 to consolidate/clean/switch to proof relying on supporting hyperplanes of Bayes Risks
%    \jessie{Skip and go to orange for now.}
%  For the reverse implication, assume that $\risk{L} = \risk{\ell}$.
%  In what follows, we implicitly work in the affine hull of $\simplex$ by taking an appropriate invertible linear projection to an affine subspace $A : \reals^n_+ \to \reals^{n-1}_+$, so that interiors are well-defined, and $\risk{\ell}$ may be differentiable on the interior of $\simplex$.
%  (For intuition, one can consider the relative interiors in $\reals^\Y_+$, but we need to consider interiors in order to apply previous results.)
%  Since $\ell$ is discrete, $-\risk{\ell}$ is polyhedral as the pointwise maximum of a finite set of linear functions.
%  \raft{I believe we can remove mentions of power diagrams here; let's give it a try.  I don't think we need to project yet, given what we say next.}
%  The projection of its epigraph $E_{\ell}$ onto $\simplex$ forms a power diagram by Theorem~\ref{thm:aurenhammer}, whose cells are full-dimensional (in $n-1$ dimensions) and correspond to the level sets $\gamma_r$ of $\gamma = \prop{\ell}$. \jessiet{Don't need Theorem 1; can use Lemma~\ref{lem:finite-full-dim} instead. -21 Mar 21}

% commented out 10 May to consolidate/clean
%  \raft{Good start.  Let's take it a bit slower.  I was thinking you start with the facets, which by definition have interiors.  Then pick a point *of the graph*, one on the interior of each facet.  Then observe it must come from some $r$ and $p$, i.e., $(p,\inprod{p}{\ell(r)})$.  Now you can define $S$ (all the $r$'s you just gathered).  The embedding comes next, using the Bayes risk matching to get that same graph point written as $(p,\inprod{p}{L(u)})$: define $\varphi(r) = u$ for that $r$ and $u$.  And so on.}
%  \proposedadd{Take $S$ as follows: for each facet $F$ (of which there are finitely many) of the epigraph of $\ell$, $E_{\ell}$, there is at least one $r \in \R$ such that $(p, \inprod{p}{\ell(r)})$ supports $E_\ell$ on the facet $F$.
%  Form $S$ by taking one such report for each facet.}
%  For each $r\in\R\proposedadd{S}$, let $p_r$ be a distribution in the interior of $\gamma_r$, and let $u_r \in \Gamma(p)$.
%  Observe that, by definition of the Bayes risk and $\Gamma$, for all $v\in\reals^d$ the hyperplane $v \mapsto \inprod{v}{-L(u_r)}$ supports the epigraph $E_L$ of $-\risk{L}$ at the point $(p,-\inprod{p}{L(v)})$ if and only if $v\in\Gamma(p)$.
%  Thus, the hyperplane $v \mapsto \inprod{v}{-L(u_r)}$ supports $E_L = E_\ell$ at the point $(p_r,-\inprod{p_r}{L(u_r)})$, and thus does so at the entire facet $\{(p,-\inprod{p}{L(u_r)}) : p\in\gamma_r\}$; by the above, $u_r \in \Gamma(p)$ for all such distributions as well.
%  We conclude that $u_r \in \Gamma(p) \iff p \in \gamma_r \iff r \in \gamma(p)$, satisfying condition~\eqref{eq:embed-loss} for $\varphi : r \mapsto u_r$.
%  To see that the loss values match, we merely note that the supporting hyperplanes to the facets of $E_L$ and $E_\ell$ are the same, and the loss values are uniquely determined by the supporting hyperplane.
%  In particular, if $h$ supports the facet corresponding to $\gamma_r$, we have $\ell(r)_y = L(u_r)_y = h(\delta_y)$, where $\delta_y$ is the point distribution on outcome $y$.

	\jessiet{Not sure how much this matters, but technically, the $\Sc$ here might be different from the $\Sc$ in the forward implication, and might even be different cardinalities. Not sure if we want to change the notation to reflect this, or if that would just add confusion.}
	For the reverse implication, assume $\risk{L} = \risk{\ell}$, which are polyhedral functions as $\ell$ is discrete.  
	Construct a minimum representative set $\Sc$ via a bijection from the facets of $E_\ell$ as in the proof of Proposition~\ref{prop:SL-minimum}, and $\U$ similarly from $E_L$, with the embedding $\varphi : r \mapsto u_r$ if the hyperplanes $p \mapsto (p, \inprod{p}{\ell(r)})$ and $p \mapsto (p, \inprod{p}{L(u_r)})$ are equal.
	Such an embedding is well-defined as it is the composition of the bijections from reports in $\Sc$ and $\U$ to facets.
	%Since $\risk{\ell}$ is polyhedral, the epigraph of its negation, $E_\ell$, has a finite set of facets, which by definition have well-defined interiors.  For each facet, there must then be some distribution $p_r$ and report $r \in \R$ such that the hyperplane $p_r \mapsto (p_r, \inprod{p_r}{-\ell(r)})$ uniquely supports $E_\ell$.  Form $\Sc$ from the set of such reports $r$, iterating over each facet.  We have $\Sc$ representative for $\ell$ since the union of projected epigraph facets equals the simplex, and these facets and their projections are closed. Therefore, every $p \in \simplex$ is contained in some facet of the epigraph, and thus there is some $r \in \Sc$ so that $r \in \gamma(p)$.
%As the risks are equal, the same is true for $E_L$, e.g., for each facet, there is some $u_r$ so that $p_r \mapsto (p_r, \inprod{p_r}{L(u_r)})$, and one can construct the embedding $\varphi: r \mapsto u_r$ for each $r \in \Sc$.

We match optimality of embedded reports by construction, since the optimal reports support the epigraph of negative Bayes risk on exactly the same distributions, meaning a report and its embedding are optimal at exactly the same distributions.
To verify matching losses, consider that the hyperplanes $p \mapsto (p, \inprod{p}{\ell(r)})$ and $p \mapsto (p, \inprod{p}{L(\varphi(r))})$ are exactly equal since they are linear functions matching on a set of positive measure relative to the simplex, and must be equal everywhere.
This implies $\ell(r) = L(\varphi(r))$ for all $r \in \Sc$ such that the hyperplanes support $E_\ell$ and $E_L$ on a facet.
Since $\Sc$ and $\U$ are minimum representative sets, this holds for all $r \in \Sc$.
%To verify matching losses, observe $\risk \ell (\delta_y) = \risk L (\delta_y)$ for each $y \in \Y$, and these risks are attained by some report $r$ and its embedding $u_r$, respectively.
%\jessie{This feels like it's missing a detail about how $\Sc$ is minimum and how level sets are full-dimensional.}
\end{proof}

@@

\raft{I wonder if we should demote this to a corollary.  We can write the proof of the reverse implication in Proposition~\ref{prop:embed-bayes-risks} like this: first show that if $\risk{L}$ is polyhedral, there is a finite $\U$ which is representative for $L$.  Then show that if $\risk{\ell} = \risk{L}$ then there is some $\Sc$ that is bijective to the particular $\U$ we constructed.  I'm not sure how this will go, but perhaps worth a try, since it would avoid the proof duplication too!}
\begin{proposition}\label{prop:embed-risk-poly}
  A loss $L$ embeds a discrete loss $\ell : \R \to \reals^\Y_+$ if and only if $\risk{L}$ is polyhedral.
\end{proposition}
\begin{proof}
  If $L$ embeds a discrete $\ell$, Proposition~\ref{prop:embed-bayes-risks} gives us $\risk{L} = \risk{\ell}$ and polyhedral.
  For the converse, let $\risk{L}$ be polyhedral. %; we again examine the proof of Proposition~\ref{prop:embed-bayes-risks}.
  Form the finite set $\Sc$ as the proof of Proposition~\ref{prop:SL-minimum}.
  %For each facet of $E_L$, which is full-dimensional, there must be some interior distribution $p$ such that $p \mapsto (p, \inprod{p}{L(u)})$ uniquely supports $E_L$, and such a $u$ must exist or else the Bayes risk is not attained.  Take $\Sc$ to be the collection of such reports $u$, and observe $\Sc$ is representative for $L$.  
  By Proposition~\ref{prop:representative-embeds-restriction}, $L$ embeds $L|_\Sc$, which is finite by construction.
%  The projection of the epigraph of $\risk{L}$ onto $\simplex$ forms a power diagram by Theorem~\ref{thm:aurenhammer} with finitely many cells $C_1,\ldots,C_k$, which we can index by $\R := \{1,\ldots,k\}$.
%  Defining the property $\gamma:\simplex\toto\R$ by $\gamma_r = C_r$ for $r\in\R$, we see that the same construction gives us points $u_r \in\reals^d$ such that $u_r \in \Gamma(p) \iff r \in \gamma(p)$.
%  Define $\ell:\R\to\reals^\Y_+$ by $\ell(r) = L(u_r)$, the same proof shows that $L$ embeds $\ell$.
\end{proof}

This insight lets us delve further into understanding characteristics of polyhedral surrogates.

@@

>>>>

\subsection{Structure of polyhedral Bayes risks}

\raf{Theme: expanding our results from polyhedral losses to polyhedral Bayes risks. We can start by observing that the latter is a weaker condition.}

\jessie{Lemma X}
\begin{lemma}\label{lem:X}
  Let $L: \R \to \reals^\Y_+$ be a minimizable loss with a polyhedral Bayes risk $\risk L$.
  Then $L$ has a finite minimum representative set.
  Furthermore, letting $\Gamma = \prop{L}$, there exist finite sets
  $\V \subseteq \reals^\Y_+$ and
  $\Theta = \{\theta_v \subseteq \simplex \mid v\in\V\}$,
  both uniquely determined by $\risk{L}$ alone,
  such that
  \begin{enumerate}
  % \item There exists $\V = \V(\risk{L}) \subseteq \reals^\Y_+$, uniquely determined by $\risk{L}$ alone, such that $\V = L(\R^*)$.\label{item:X-V}
  \item A set $\R'\subseteq\R$ is representative if and only if $\V \subseteq L(\R')$.\label{item:X-rep-V}
  \item A set $\R'\subseteq\R$ is minimum representative if and only if $L(\R') = \V$.\label{item:X-min-V}
  \item A set $\R'\subseteq\R$ is representative if and only if $\Theta \subseteq \{\Gamma_r \mid r \in \R'\}$.\label{item:X-rep-Theta}
  \item A set $\R'\subseteq\R$ is minimum representative if and only if $\{\Gamma_r \mid r \in \R'\} = \Theta$.\label{item:X-min-Theta}
  \item The set of full-dimensional level sets of $\Gamma$ is exactly $\Theta$.\label{item:X-full-dim}
  \item For any $r \in \R$, there exists $\theta \in \Theta$ such that $\Gamma_r \subseteq \theta$.\label{item:X-redundant}
  \end{enumerate}
\end{lemma}

\raf{Make a corollary.  THus we'll focus on finite rep sets.}

\begin{proposition}\label{prop:SL-minimum}
  A minimizable loss $L: \R \to \reals^\Y_+$ has a polyhedral Bayes risk if and only if it has a finite representive set.
  % Moreover, for any minimum representative set $\Sc$ for $L$, there is a bijection between $\Sc$ and the facets of the epigraph of $-\risk{L}$.
  % \raft{Jessie: prove :-].  Note: min rep $\implies$ poly risk is immediate from Lemma~\ref{lem:loss-restrict}.  Move the construction into the proof. \jessie{Done; ready for a look}}
\end{proposition}
\begin{proof}
  A finite representative implies a polyhedral Bayes risk by Lemma~\ref{lem:loss-restrict}.
  Lemma~\ref{lem:X} gives the converse.
\end{proof}

@@

@@

\begin{proposition}\label{prop:embed-bayes-risks}
  Let discrete loss $\ell$ and minimizable loss $L$ be given.
  Then $L$ embeds $\ell$ if and only if $\risk{L}=\risk{\ell}$.
  % Let discrete loss $\ell:\R\to\reals^Y$ be given.
  % Then $L:\reals^d\to\reals^\Y$ embeds $\ell$ if and only if $\risk{L}=\risk{\ell}$.
  % A loss $L$ embeds a discrete loss $\ell$ if and only if $\risk{L}=\risk{\ell}$.
\end{proposition}
\begin{proof}
  Define $\Gamma = \prop{L}$ and $\gamma = \prop{\ell}$.
  Suppose $L$ embeds $\ell$, so we have some $\Sc\subseteq \R$ which is representative for $\ell$ and an embedding $\varphi:\Sc\to\reals^d$; take $\U := \varphi(\Sc)$.
  Since $\Sc$ is representative for $\ell$, by embedding condition (ii) we have $\{\gamma_s \mid s\in\Sc\} = \{\Gamma_u \mid u\in\U\}$, so $\U$ is representative for $L$.
  % for all $p\in\simplex$ we have some $r \in \Sc \cap \gamma(p)$ and thus we have $\varphi(r) \in \U \cap \Gamma(p)$.
  % In particular, $\U \cap \Gamma(p)\neq \emptyset$ for all $p\in\simplex$, 
  By Lemma~\ref{lem:loss-restrict}, we have $\risk{\ell} = \risk{\ell|_{\Sc}}$ and $\risk{L} = \risk{L|_{\U}}$.
  As $L(\varphi(\cdot)) = \ell(\cdot)$ by embedding condition (i), for all $p\in\simplex$ we have
  \begin{equation*}
    \risk{\ell}(p) = \risk{\ell|_\Sc}(p) = \min_{r \in \Sc}\inprod{p}{\ell(r)} = \min_{r \in \Sc}\inprod{p}{L(\varphi(r))} = \min_{u \in \U}\inprod{p}{L(u)} = \risk{L|_\U}(p) = \risk{L}(p)~.
  \end{equation*}

	For the reverse implication, assume $\risk{L} = \risk{\ell}$, which are polyhedral functions as $\ell$ is discrete.
  From Lemma~\ref{lem:X}(\ref{item:X-min-V}), we have some set $\V\subseteq\reals^\Y_+$ and minimum representative sets $\R^* \subseteq \R$ and $\Sc \subseteq \U$, for $\ell$ and $L$ respectively, such that $\ell(\R^*) = \V = L(\Sc)$.
  As $\R^*$ and $\Sc$ are miniumum, they cannot repeat loss vectors, and thus $|\R^*|=|\ell(\R^*)|$ and $|L(\Sc)|=\Sc$.
  We conclude that $\R^*$ and $\Sc$ are both in bijection with $\V$.
  The map $\varphi :\R^* \to \reals^d$, given by $\varphi(r) = s \in \Sc$ where $\ell(r) = L(s)$, is therefore well-defined.
  Condition (i) of an embedding is immediate.
  From Proposition~\ref{prop:representative-embeds-restriction}, $\ell$ embeds $\ell|_{\R^*}$ and $L$ embeds $L|_\Sc$, both via the identity embedding.
  Thus for all $p\in\simplex$ and $r\in\R^*$, we have
  \begin{equation*}
    r \in \gamma(p) \iff r \in \prop{\ell|_{\R^*}}(p) \iff \varphi(r) \in \prop{L|_\Sc}(p)
    \iff \varphi(r) \in \prop{L}(p)~,
  \end{equation*}
  giving condition (ii).
\end{proof}

@@

\raf{Demote to corollary?}
\begin{proposition}\label{prop:embed-risk-poly}
  A minimizable loss $L$ embeds a discrete loss $\ell$ if and only if $\risk{L}$ is polyhedral.
\end{proposition}
\begin{proof}
  If $L$ embeds a discrete $\ell$, Proposition~\ref{prop:embed-bayes-risks} gives us $\risk{L} = \risk{\ell}$ and therefore $\risk{L}$ is polyhedral.
  For the converse, let $\risk{L}$ be polyhedral. %; we again examine the proof of Proposition~\ref{prop:embed-bayes-risks}.
  Lemma~\ref{lem:X} gives a finite representative set $\Sc$, and $L$ embeds $L|_\Sc$ by Proposition~\ref{prop:representative-embeds-restriction}.
\end{proof}

\raf{Moved; reorganize}

@@

\subsection{A property elicitation perspective}
